### 2014-11-04 paper/paper.tex

\documentclass{sig-alternate}

%% set up bibliography
\usepackage[backend=bibtex,doi=false,isbn=false,url=false]{biblatex}
\bibliography{biblio}  % sigproc.bib is the name of the Bibliography in this case
%\bibliographystyle{acm}

\begin{document}

\title{State of the art in {\ttlit Sentiment Analysis}}

%\subtitle{}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{5} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
C. B. Hotz-Behofsits\\
       \affaddr{0929002}\\
       \affaddr{christian.hotz-behofsits@tuwien.ac.at}
% 2nd. author
\alignauthor
Thomas Schmidtleitner\\
       \affaddr{1025525}\\
       \affaddr{e1025525@student.tuwien.ac.at}
% 3rd. author
\alignauthor
Dominik Pichler\\
       \affaddr{1026045}\\
       \affaddr{dominik.pichler@aon.at}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor
Matthias Reisinger\\
       \affaddr{1025631}\\
       \affaddr{matthias.reisinger@web0.at}
% 5th. author
\alignauthor
Florian Taus\\
       \affaddr{0627918}\\
       \affaddr{florian.taus@hotmail.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}

\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\input{chapters/1_introduction.tex}
\input{chapters/2_methods.tex}
\input{chapters/3_conclusions.tex}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}

\printbibliography

\end{document}