### 2014-11-27 paper/chapters/3_preprocessing.tex

\section{Preprocessing}
The Data-Preprocessing-Process is an essential part of sentiment-Analysis. Its goal is to prepare data
for the sentiment analysis and remove unnecessary parts.\autocite{Hemalatha2012}

Unnecessary Parts are:\autocite{Hemalatha2012}
\begin{enumerate}
    \item Remove URLs, Special Characters
    \item Filter Unnecesarry Words
    \item Remove Retweets
\end{enumerate}

Besondere an twitter: Hashtags, Mentions, Smileys
Tokenizen